MSTK: Mesh Toolkit, v 1.7 - DRAFT

### 2010-03-20 docs/MSTK.tex
Updated version of MSTK manual with new operators, new examples and new sections

    \begin{minipage}{1.75in}
      \begin{center}
        \includegraphics[scale=0.9]{figures/repR1} \\
        Representation R1
      \end{center}
    \end{minipage}
    \begin{minipage}{1.75in}
      \begin{center}
        \includegraphics[scale=0.9]{figures/repR2} \\
        Representation R2
      \end{center}
    \end{minipage}
    \begin{minipage}{1.75in}
      \begin{center}
        \includegraphics[scale=0.9]{figures/repR4} \\
        Representation R4
      \end{center}
    \end{minipage}
  \end{center}
\end{figure}

@@

\item[]{\bf {\em RepType}}: Enumerated type describing the type
  of mesh representation.  Can be UNKNOWN\_REP, F1, F4, R1, R2, R4.
  See Appendix~\ref{app:reptypes} for schematics of these
  representations. {\em Currently only representation types F1 and
    F4 are supported}.

@@

\item[]{\bf {\em AttType}}: Enumerated type for attribute data. Can be INT, DOUBLE or POINTER
\end{description}

@@

\item[]{\bf {\em int} MESH\_InitFromFile({\em Mesh\_ptr}
    mesh, const char *filename):} Initialize or read a mesh from file
  into the given mesh object. Returns 1 if successful, 0 otherwise. It
  is possible to have a an MSTK file in the R1, R2 and R4 format into
  a mesh initialized as type F1 or F4.

@@

\item[]{\bf {\em int} MESH\_ImportFromFile({\em Mesh\_ptr}
    mesh, const char *filename, const char *format):} Import a mesh
  data from an external file format and construct MSTK mesh. Currently
  the only format supported is the
  GMV\footnote{\texttt{http://www-xdiv.lanl.gov/XCM/gmv/GMVHome.html}}
  file format. The routine imports as many attributes as it can from
  the input file. It also uses special attributes or keywords as data
  about element and node classification.  For GMV files, the routine
  uses the ``material'' data to indicate region or face classification
  depending on the dimensionality of the mesh. If ``material'' data is
  not specified, it uses the ``itetclr'' attribute to assign region or
  face classification. It also uses the ``icr'' keyword describing the
  number of constraints on a node to interpret if the node is
  classified on a model region, model face, model edge or a model
  vertex. The ``itetclr'' and ``icr'' keywords are usually present in
  meshes generated by
  LAGRIT\footnote{\texttt{http://lagrit.lanl.gov'}}.

@@

\item[]{\bf {\em void} MESH\_WriteToFile({\em Mesh\_ptr}
    mesh, const char *filename):} Save a mesh to a filename. The file
  is created if it does not exists. It is recommended that the .mstk
  extension be used for MSTK mesh files.  However, there is no such
  requirement.

@@

\item[]{\bf {\em List\_ptr} MV\_Edges({\em MVertex\_ptr} mvertex):}
List of edges connected to the vertex.

\item[]{\bf {\em List\_ptr} MV\_Faces({\em MVertex\_ptr} mvertex):}
List of faces connected to the vertex.

\item[]{\bf {\em List\_ptr} MV\_Regions({\em MVertex\_ptr} mvertex):}
List of regions connected to the vertex.
\end{description}

@@

\item[]{\bf {\em List\_ptr} ME\_Faces({\em MEdge\_ptr}
    medge):} Returns the set of faces using this edge.

\item[]{\bf {\em List\_ptr} ME\_Regions({\em MEdge\_ptr}
    medge):} Returns the set of regions using this edge.

@@

\item[]{\bf {\em List\_ptr} MF\_Vertices({\em MFace\_ptr}
    mface, {\em int} dir, {\em MVertex\_ptr} mvert):} Return the
  ordered set of the vertices of the face. The vertices are ordered in
  ccw direction while looking down the face 'normal', if 'dir' is 1
  and in the cw direction, if 'dir' is 0. If 'mvert' is specified, the
  vertex set is reordered so that it is the first vertex (This
  argument will be added soon to the function. For now, omit this
  argument). The behavior of this function can be illustrated using
  Figure~\ref{fig:face_def}. For the face shown in the figure, a
  vertex set with ccw ordering or 'dir' = 1 is ${V_0,V_1,V_2,V_3}$ and
  a vertex set with cw ordering or 'dir' = 0 is ${V_0,V_3,V_2,V_1}$.
  A vertex set with ccw ordering starting with vertex $V_2$ is
  ${V_2,V_3,V_0,V_1}$.

@@

\item[]{\bf {\em List\_ptr} MF\_Edges({\em MFace\_ptr} mface,
    {\em int} dir, {\em MVertex\_ptr} mvert):} Return the
  ordered set of edges of the face. The edges are ordered in the ccw
  while looking down the face 'normal' if dir is 1 and in the cw if
  dir is 0. If 'mvert' is specified, the edge set is reordered so that
  the first edge in the set contains this vertex. More precisely, if
  'dir' is 1, and the first edge is 'e' used in the face in direction
  'd', then {\bf ME\_Vertex(e,!d)} = mvert. With reference to
  Figure~\ref{fig:face_def}, the edges of the face in the ccw direction or
  'dir' = 1 are ${E_0,E_1,E_2,E_3}$ and in the opposite dir are
  ${E_3,E_2,E_1,E_0}$. If 'mvert' or the starting vertex is specified
  as $V_2$, the edge set in ccw direction or 'dir' = 1 is
  ${E_2,E_3,E_0,E_1}$ and in the opposite direction is
  ${E_1,E_0,E_3,E_2}$.

\item[]{\bf {\em List\_ptr} MF\_AdjFaces({\em MFace\_ptr} mface):}
List of adacent faces of a face. This operator is relevant only in
planar or surface surface meshes, i.e., for boundary faces not
connected to any regions.

@@

\item[]{\bf {\em List\_ptr} MF\_Regions({\em MFace\_ptr} mface):}
Return the set of regions connected to the face. If the face is not
used by any regions, the function returns NULL to indicate that the
set is empty. If not the set may contain one or two regions.

@@

\item[]{\bf {\em List\_ptr} MR\_Vertices({\em MRegion\_ptr}
mregion):} Returns the set of vertices of a region. For standard
elements the vertices are ordered as indicated in
Appendix~\ref{app:reg_conventions}. For non-standard elements the set
is unordered.

\item[]{\bf {\em List\_ptr} MR\_Edges({\em MRegion\_ptr} mregion):}
Return the \underline{unordered} set of edges of a region.

\item[]{\bf {\em List\_ptr} MR\_Faces({\em MRegion\_ptr} mregion):}
Returns the set of faces of a region.

@@

\item[]{\bf {\em void} MEnt\_Set\_AttVal({\em MEntity\_ptr}
    ent, {\em MAttrib\_ptr} attrib, {\em int} ival,
    {\em double} lval, {\em void} *pval):} Set the value of an
  attribute for given mesh entity. Depending on whether the attribute
  is integer, double or pointer, the appropriate entry in the argument
  list is specified.

@@

\item[]{\bf {\em int} MEnt\_Get\_AttVal({\em MEntity\_ptr}
    ent, {\em MAttrib\_ptr} int *ival, double *rval, double
    **pval):} Get the value of an attribute from an entity.  Depending
  on the attribute type, either {\bf ival}, {\bf rval} or
  {\bf pval} is returned.

@@

\item[]{\bf {\em MAttrib\_ptr} MAttrib\_New({\em Mesh\_ptr}
    mesh, {\em const char} *att\_name, {\em MAttType} att\_type,
    {\em MType} entdim):} Define a new mesh attribute. Along with
  the mesh to which this attribute is assigned ({\bf mesh}) and the
  name of the attribute ({\bf att\_name}), the type of the
  attribute ({\bf att\_type}) must be specified as {\bf INT},
  {\bf DOUBLE} or {\bf POINTER}. Also, the dimension of the entity
    for which the attribute value can be set must be specified as
    {\bf MVERTEX}, {\bf MEDGE}, {\bf MFACE}, {\bf MREGION}
    or {\bf MALLTYPE}.

@@

\# {\em {\bf NOT IMPLEMENTED}}

\# {\em {\bf VERTEX ATTRIBUTES}} \\
\# {\em {\em int} NVA - Number of Vertex attributes} \\
\# \\
\# {\em char *VA\_name\_1 - Name of first vertex attribute} \\
\# {\em {\em int} VA\_type\_1 - Type of first vertex attribute} \\
\# {\em {\em int} VA\_dim\_1 - Dimension of first vertex attribute} \\
\# \\
\# {\em char *VA\_name\_2 - Name of second vertex attribute} \\
\# {\em {\em int} VA\_type\_2 - Type of second vertex attribute} \\
\# {\em {\em int} VA\_dim\_2 - Dimension of first vertex attribute} \\
\# \\
\# . . .\\
\# \\
\# {\em char *VA\_name\_NVA - Name of NVA'th vertex attribute} \\
\# {\em {\em int} VA\_type\_NVA - Type of NVA'th vertex attribute} \\
\# {\em {\em int} VA\_dim\_NVA - Dimension of NVA'th vertex attribute} \\
\# \\
\# {\em VA\_type can be 1 (int), 2 (double), 3 (string)} \\
\# {\em VA\_dim = 1 for scalars, VA\_dim = length of vector for vector} \\
\# {\em VA\_dim can only be 1 when VA\_type is string}
\vspace{1ex}

NVA \hspace{0.5ex} \\
VA\_name\_1 \hspace{0.5ex} VA\_type\_1 \hspace{0.5ex} VA\_dim\_1 \\ 
VA\_name\_2 \hspace{0.5ex} VA\_type\_2 \hspace{0.5ex} VA\_dim\_2 \\
. \\
VA\_dim\_NVA \hspace{0.5ex} VA\_type\_NVA \hspace{0.5ex} VA\_dim\_2  
\vspace{1ex}

\# {\em For each vertex attribute record, set of attribute values} \\
\# {\em E.G., there are 3 attributes for each vertex:} \\
\# {\em a scalar int, a vector of 3 doubles and a string}

VA\_{\em int} \hspace{0.5ex} VA\_double\_1 \hspace{0.5ex} VA\_double\_2 \hspace{0.5ex} VA\_double\_3 \hspace{0.5ex} VA\_string \\
VA\_{\em int} \hspace{0.5ex} VA\_double\_1 \hspace{0.5ex} VA\_double\_2 \hspace{0.5ex} VA\_double\_3 \hspace{0.5ex} VA\_string \\
. \\
. \\
. \\
\# {\em Repeated NV times}
\vspace{2ex}

\newpage
\# {\em {\bf EDGE ATTRIBUTES}} \\
\# {\em Similar to Vertex attribute description}
\vspace{1ex}

NEA \hspace{0.5ex} \\
EA\_name\_1 \hspace{0.5ex} EA\_type\_1 \hspace{0.5ex} EA\_dim\_1 \\ 
EA\_name\_2 \hspace{0.5ex} EA\_type\_2 \hspace{0.5ex} EA\_dim\_2 \\
. \\
EA\_dim\_NEA \hspace{0.5ex} EA\_type\_NEA \hspace{0.5ex} EA\_dim\_2  
\vspace{1ex}

\# {\em For each edge attribute record, set of attribute values} \\
\# {\em E.G., a scalar int, a scalar double and a string}
\vspace{1ex}

EA\_{\em int} \hspace{0.5ex} EA\_double \hspace{0.5ex} EA\_string \\
EA\_{\em int} \hspace{0.5ex} EA\_double \hspace{0.5ex} EA\_string \\
. . . \\
\# {\em Repeated NE times}
\vspace{2ex}

\# {\em {\bf FACE ATTRIBUTES}} \\
\# {\em Similar to Vertex attribute description}
\vspace{1ex}

NFA \hspace{0.5ex} \\
FA\_name\_1 \hspace{0.5ex} FA\_type\_1 \hspace{0.5ex} FA\_dim\_1 \\ 
FA\_name\_2 \hspace{0.5ex} FA\_type\_2 \hspace{0.5ex} FA\_dim\_2 \\
. \\
FA\_dim\_NFA \hspace{0.5ex} FA\_type\_NEA \hspace{0.5ex} FA\_dim\_2  
\vspace{1ex}

\# {\em For each face attribute record, set of attribute values} \\
\# {\em E.G., a vector of 2 doubles and a string}
\vspace{1ex}

FA\_double\_1 \hspace{0.5ex} FA\_double\_2 \\
FA\_double\_1 \hspace{0.5ex} FA\_double\_2 \\
. . . \\
\# {\em Repeated NF times}
\vspace{2ex}

\newpage
\# {\em {\bf REGION ATTRIBUTES}} \\
\# {\em Similar to Vertex attribute description}
\vspace{1ex}

NRA \hspace{0.5ex} \\
RA\_name\_1 \hspace{0.5ex} RA\_type\_1 \hspace{0.5ex} RA\_dim\_1 \\ 
RA\_name\_2 \hspace{0.5ex} RA\_type\_2 \hspace{0.5ex} RA\_dim\_2 \\
. \\
RA\_dim\_NRA \hspace{0.5ex} RA\_type\_NRA \hspace{0.5ex} RA\_dim\_NRA 
\vspace{1ex}

\# {\em For each Region attribute record, set of attribute values} \\
\# {\em E.G., a vector of 3 ints}
\vspace{1ex}

RA\_int\_1 \hspace{0.5ex} RA\_int\_2 \hspace{0.5ex} RA\_int\_3 \\
RA\_int\_1 \hspace{0.5ex} RA\_int\_2 \hspace{0.5ex} RA\_int\_3 \\
. . .\\
\# {\em Repeated NR times}
\vspace{2ex}

@@

@@

@@

 mesh = MESH_New(UNKNOWN_REP);
 ok = MESH_InitFromFile(mesh,meshname);
 if (!ok) {
   fprintf(stderr,"Cannot file input file %s\n\n\n",meshname);
   exit(-1);
 }

@@

>>>>

    \begin{minipage}{1.75in}
      \begin{center}
        \includegraphics[scale=0.9]{figures/repR1} \\
        Representation R1
      \end{center}
    \end{minipage}
    \begin{minipage}{1.75in}
      \begin{center}
        \includegraphics[scale=0.9]{figures/repR2} \\
        Representation R2
      \end{center}
    \end{minipage}
    \begin{minipage}{1.75in}
      \begin{center}
        \includegraphics[scale=0.9]{figures/repR4} \\
        Representation R4
      \end{center}
    \end{minipage}
  \end{center}
  \caption{Mesh representations in MSTK}
  \label{fig:reptypes}
\end{figure}

@@

\item[]{\bf {\em RepType}}: Enumerated type describing the type
  of mesh representation.  Can be UNKNOWN\_REP, F1, F4, R1, R2, R4.
  See Appendix~\ref{fig:reptypes} for schematics of these
  representations. {\em Currently only representation types F1 and
    F4 are supported}.

@@

\item[]{\bf {\em AttType}}: Enumerated type for attribute data. Can be INT, DOUBLE, POINTER, VECTOR or TENSOR.
\end{description}

@@

\item[]{\bf {\em int} MESH\_InitFromFile({\em Mesh\_ptr} mesh, const
    char *filename):} Initialize or read a mesh from a file in the
  MSTK format into the given mesh object. Returns 1 if successful, 0
  otherwise. It is possible to have a an MSTK file in the R1, R2 and
  R4 format into a mesh initialized as type F1 or F4. This routine
  imports any attributes that are into the mesh.

@@

\item[]{\bf {\em int} MESH\_ImportFromFile({\em Mesh\_ptr}
    mesh, const char *filename, const char *format):} Import a mesh
  data from an external file format and construct MSTK mesh. Currently
  the only format supported is the
  GMV\footnote{\texttt{http://www-xdiv.lanl.gov/XCM/gmv/GMVHome.html}}
  file format. The routine imports as many attributes as it can from
  the input file. It also uses special attributes or keywords as data
  about element and node classification.  For GMV files, the routine
  uses the ``material'' data to indicate region or face classification
  depending on the dimensionality of the mesh. If ``material'' data is
  not specified, it uses the ``itetclr'' attribute to assign region or
  face classification. It also uses the ``icr'' keyword describing the
  number of constraints on a node to interpret if the node is
  classified on a model region, model face, model edge or a model
  vertex. The ``itetclr'' and ``icr'' keywords are usually present in
  meshes generated by
  LAGRIT\footnote{\texttt{http://lagrit.lanl.gov}}.

@@

  \item[]{\bf {\em void} MESH\_WriteToFile({\em Mesh\_ptr} mesh, const
      char *filename):} Save a mesh to a filename. The file is created
    if it does not exists. It is recommended that the .mstk extension
    be used for MSTK mesh files.  However, there is no such
    requirement. The routine will write out the mesh and any
    attributes attached to the mesh except attributes of the type
    POINTER.

@@

\item[]{\bf {\em List\_ptr} MV\_Edges({\em MVertex\_ptr} mvertex):}
List of edges connected to the vertex. {\bf The list returned by this
  operator must be deleted by the calling application using {\em List\_Delete}}.

\item[]{\bf {\em List\_ptr} MV\_Faces({\em MVertex\_ptr} mvertex):}
List of faces connected to the vertex. {\bf The list returned by this
  operator must be deleted by the calling application using {\em List\_Delete}}.

\item[]{\bf {\em List\_ptr} MV\_Regions({\em MVertex\_ptr} mvertex):}
List of regions connected to the vertex. {\bf The list returned by this
  operator must be deleted by the calling application using {\em List\_Delete}}.
\end{description}

@@

\item[]{\bf {\em List\_ptr} ME\_Faces({\em MEdge\_ptr}
    medge):} Returns the set of faces using this edge. {\bf The list returned by this
  operator must be deleted by the calling application using {\em List\_Delete}}.

\item[]{\bf {\em List\_ptr} ME\_Regions({\em MEdge\_ptr}
    medge):} Returns the set of regions using this edge. {\bf The list returned by this
  operator must be deleted by the calling application using {\em List\_Delete}}.

@@

\item[]{\bf {\em List\_ptr} MF\_Vertices({\em MFace\_ptr}
    mface, {\em int} dir, {\em MVertex\_ptr} mvert):} Return the
  ordered set of the vertices of the face. The vertices are ordered in
  ccw direction while looking down the face 'normal', if 'dir' is 1
  and in the cw direction, if 'dir' is 0. If 'mvert' is specified, the
  vertex set is reordered so that it is the first vertex (This
  argument will be added soon to the function. For now, omit this
  argument). The behavior of this function can be illustrated using
  Figure~\ref{fig:face_def}. For the face shown in the figure, a
  vertex set with ccw ordering or 'dir' = 1 is ${V_0,V_1,V_2,V_3}$ and
  a vertex set with cw ordering or 'dir' = 0 is ${V_0,V_3,V_2,V_1}$.
  A vertex set with ccw ordering starting with vertex $V_2$ is
  ${V_2,V_3,V_0,V_1}$.  {\bf The list returned by this
  operator must be deleted by the calling application using {\em List\_Delete}}.

@@

\item[]{\bf {\em List\_ptr} MF\_Edges({\em MFace\_ptr} mface,
    {\em int} dir, {\em MVertex\_ptr} mvert):} Return the
  ordered set of edges of the face. The edges are ordered in the ccw
  while looking down the face 'normal' if dir is 1 and in the cw if
  dir is 0. If 'mvert' is specified, the edge set is reordered so that
  the first edge in the set contains this vertex. More precisely, if
  'dir' is 1, and the first edge is 'e' used in the face in direction
  'd', then {\bf ME\_Vertex(e,!d)} = mvert. With reference to
  Figure~\ref{fig:face_def}, the edges of the face in the ccw direction or
  'dir' = 1 are ${E_0,E_1,E_2,E_3}$ and in the opposite dir are
  ${E_3,E_2,E_1,E_0}$. If 'mvert' or the starting vertex is specified
  as $V_2$, the edge set in ccw direction or 'dir' = 1 is
  ${E_2,E_3,E_0,E_1}$ and in the opposite direction is
  ${E_1,E_0,E_3,E_2}$.  {\bf The list returned by this
  operator must be deleted by the calling application using {\em List\_Delete}}.

\item[]{\bf {\em List\_ptr} MF\_AdjFaces({\em MFace\_ptr} mface):}
List of adacent faces of a face. This operator is relevant only in
planar or surface surface meshes, i.e., for boundary faces not
connected to any regions.  {\bf The list returned by this
  operator must be deleted by the calling application using {\em List\_Delete}}.

@@

\item[]{\bf {\em List\_ptr} MF\_Regions({\em MFace\_ptr} mface):}
Return the set of regions connected to the face. If the face is not
used by any regions, the function returns NULL to indicate that the
set is empty. If not the set may contain one or two regions.  {\bf The list returned by this
  operator must be deleted by the calling application using {\em List\_Delete}}.

@@

\item[]{\bf {\em List\_ptr} MR\_Vertices({\em MRegion\_ptr}
mregion):} Returns the set of vertices of a region. For standard
elements the vertices are ordered as indicated in
Appendix~\ref{app:reg_conventions}. For non-standard elements the set
is unordered. {\bf The list returned by this
  operator must be deleted by the calling application using {\em List\_Delete}}.

\item[]{\bf {\em List\_ptr} MR\_Edges({\em MRegion\_ptr} mregion):}
Return the \underline{unordered} set of edges of a region.  {\bf The list returned by this
  operator must be deleted by the calling application using {\em List\_Delete}}.

\item[]{\bf {\em List\_ptr} MR\_Faces({\em MRegion\_ptr} mregion):}
Returns the set of faces of a region. {\bf The list returned by this
  operator must be deleted by the calling application using {\em List\_Delete}}.

@@

\item[]{\bf {\em void} MEnt\_Set\_AttVal({\em MEntity\_ptr} ent, {\em
      MAttrib\_ptr} attrib, {\em int} ival, {\em double} lval, {\em
      void} *pval):} Set the value of an attribute for given mesh
  entity. Depending on the attribute type, different arguments of this
  function are relevant. If the attribute type is:

\begin{itemize}
  \item {\bf INT} -- specify the integer value in {\bf ival},
  \item {\bf DOUBLE} -- specify the the real value in {\bf rval},
  \item {\bf VECTOR} -- specify a pointer to an array of
    doubles in {\bf pval}\footnote{Remember to not free the array specified in {\bf pval} without telling the attribute system because MSTK stores the pointer to the array directly and does not make a copy of the array.} - the number of doubles should equal the number components specified in the definition of the VECTOR attribute,
  \item {\bf TENSOR} -- specify a pointer
    to an array of doubles in {\bf pval} - the number of doubles should equal the number of components specified in the definition of the TENSOR attribute,
  \item {\bf POINTER} -- specify the pointer value in {\bf pval}.
  \end{itemize}

@@

\item[]{\bf {\em int} MEnt\_Get\_AttVal({\em MEntity\_ptr} ent, {\em
      MAttrib\_ptr} int *ival, double *rval, double **pval):} Get the
  value of an attribute from an entity.  Depending on the attribute
  type, different arguments of this function are relevant. If the
  attribute type is:

  \begin{itemize}
  \item {\bf INT} -- {\bf ival} contains the integer value,
  \item {\bf DOUBLE} -- {\bf rval} contains the real value,
  \item {\bf VECTOR} -- {\bf pval} contains a pointer to an array of doubles,
  \item {\bf TENSOR} -- {\bf pval} contains a pointer
    to an array of doubles,
  \item {\bf POINTER} -- {\bf pval} conntains the pointer value.
  \end{itemize}

@@

\item[]{\bf {\em MAttrib\_ptr} MAttrib\_New({\em Mesh\_ptr} mesh, {\em
      const char} *att\_name, {\em MAttType} att\_type, {\em MType}
    entdim, $<${\em int ncomponents}$>$):} Define a new mesh
  attribute. Along with the mesh to which this attribute is assigned
  ({\bf mesh}) and the name of the attribute ({\bf att\_name}), the
  type of the attribute ({\bf att\_type}) must be specified as {\bf
    INT}, {\bf DOUBLE}, {\bf VECTOR}, {\bf TENSOR} or {\bf
    POINTER}. Also, the dimension of the entity for which the
  attribute value can be set must be specified as {\bf MVERTEX}, {\bf
    MEDGE}, {\bf MFACE}, {\bf MREGION} or {\bf MALLTYPE}. The last
  argument {\em {\bf ncomponents}} is an optional argument for attributes of
  type {\bf INT}, {\bf DOUBLE} or {\bf POINTER} but is required for
  attributes of type {\bf VECTOR} and {\bf TENSOR}.

@@

\# {\em {\bf ATTRIBUTES}} \\
\# \\
\# {\em char *Attrib\_name\_1 - Name of first attribute} \\
\# {\em int Attrib\_type\_1 - Type of  attribute (INT, DOUBLE, VECTOR, TENSOR} \\
\# {\em int Attrib\_len\_1 - 0 (scalars), number of components (vectors), number of components (tensors)} \\
\# {\em char *Attrib\_ent\_1 - Type of mesh entity attribute is applicable to (VERTEX, EDGE, FACE, REGION, ALLTYPES)} \\
\vspace{1ex}
\# {\em int Nent - Number entities for which this attribute is specified - Useful if an attribute is to be specified for only a small subset of mesh entities, e.g. boundary conditions} \\
\# {\em Entity\_dim\_1 \hspace{0.5ex} Entity\_ID\_1 \hspace{0.5ex} Attrib\_components} \\ 
\# {\em Entity\_dim\_2 \hspace{0.5ex} Entity\_ID\_2 \hspace{0.5ex} Attrib\_components} \\ 
\# {\em Entity\_dim\_3 \hspace{0.5ex} Entity\_ID\_3 \hspace{0.5ex} Attrib\_components} \\ 
.\\
.\\
.\\
\# {\em Entity\_dim\_n \hspace{0.5ex} Entity\_ID\_n \hspace{0.5ex} Attrib\_components}
\vspace{1ex}

. \\
. \\
. \\

\# {\em char *Attrib\_name\_N - Name of first attribute} \\
\# {\em int Attrib\_type\_N - Type of  attribute (INT, DOUBLE, VECTOR, TENSOR)} \\
\# {\em int Attrib\_len\_N - 0 (scalars), number of components (vectors), number of components (tensors)} \\
\# {\em char *Attrib\_ent\_N - Type of mesh entity attribute is applicable to (VERTEX, EDGE, FACE, REGION, ALLTYPES)} \\
\vspace{1ex}
\# {\em int Nent - Number entities for which this attribute is specified} \\
\# {\em Entity\_dim\_1 \hspace{0.5ex} Entity\_ID\_1 \hspace{0.5ex} Attrib\_components} \\ 
\# {\em Entity\_dim\_2 \hspace{0.5ex} Entity\_ID\_2 \hspace{0.5ex} Attrib\_components} \\ 
\# {\em Entity\_dim\_3 \hspace{0.5ex} Entity\_ID\_3 \hspace{0.5ex} Attrib\_components} \\ 
.\\
.\\
.\\
\# {\em Entity\_dim\_n \hspace{0.5ex} Entity\_ID\_n \hspace{0.5ex} Attrib\_components}
\vspace{1ex}
\vspace{2ex}

@@

\section{Example Mesh and Files}

In this section we show a simple square geometric model and an example
2D mesh of that model comprised of triangles, quads (convex and
non-convex) and polygons. Assume that the mesh vertices have a
velocity associated with them and mesh faces have a symmetric
second-order diffusivity tensor associated with them. We then
present the MSTK files describing the mesh in the F1 and R1 formats.

\begin{figure}[!h]
  \begin{center}
    \begin{minipage}{0.5\textwidth}
      \begin{center}
        \includegraphics[scale=0.85]{figures/examplemodel}
      \end{center}
    \end{minipage}%
    \begin{minipage}{0.5\textwidth}
      \begin{center}
        \includegraphics[scale=0.85]{figures/examplemesh}
      \end{center}
    \end{minipage}
  \end{center}
  \caption{Geometric model and a two-dimensional mesh of mixed elements of the model.}
\end{figure}

File in F1 format:
\begin{verbatim}
MSTK 1.0
F1 18 31 14 0
vertices
0.00 0.00 0.00  0 1
0.33 0.00 0.00  1 1
0.66 0.00 0.00  1 1
1.00 0.00 0.00  0 2
0.00 0.33 0.00  1 4
0.66 0.66 0.00  2 1
0.66 1.00 0.00  1 3
0.16 0.84 0.00  2 1
0.00 0.66 0.00  1 4
0.16 0.33 0.00  2 1
0.50 0.33 0.00  2 1
0.40 0.84 0.00  2 1
1.00 0.66 0.00  1 2
0.00 1.00 0.00  0 4
1.00 1.00 0.00  0 3
0.66 0.50 0.00  2 1
0.33 1.00 0.00  1 3
1.00 0.33 0.00  1 2
edges 
2 1    1 1
2 3    1 1
3 4    1 1
4 18   1 2
2 11   2 1
15 7   1 3
14 8   2 1
18 16  2 1
12 16  2 1
10 5   2 1
11 16  2 1
15 13  1 2
14 17  1 3
17 7   1 3
8  9   2 1
18 13  1 2
14 9   1 4
10 11  2 1
10 2   2 1
9 10   2 1
16 6   2 1
6 13   2 1
7 6    2 1
17 12  2 1
8 17   2 1
3 11   2 1
11 18  2 1
1 5    1 4
5 9    1 4
7 12   2 1
8 12   2 1
faces edge
4  3 4 -27 -26            2 1
6  18 11 -9 -31 15 20     2 1
3  25 -13 7               2 1
3  31 -24 -25             2 1
3  2 26 -5                2 1
4  -28 -1 -19 10          2 1
4  16 -22 -21 -8          2 1
4  22 -12 6 23            2 1
3  -30 -14 24             2 1
3  -10 -20 -29            2 1
3  19 5 -18               2 1
4  9 21 -23 30            2 1
3  -15 -7 17              2 1
3  27 8 -11               2 1
attributes
velocity
VECTOR
3
MVERTEX
18
0 1    0.0  0.0  0.0
0 2    3.3  0.0  0.0
0 3    6.6  0.0  0.0
0 4   10.0  0.0  0.0
0 5    0.0  3.3  0.0
0 6    6.6  6.6  0.0
0 7    6.6 10.0  0.0
0 8    1.6  8.4  0.0
0 9    0.0  6.6  0.0
0 10   1.6  3.3  0.0
0 11   5.0  3.3  0.0
0 12   4.0  6.6  0.0
0 13  10.0  6.6  0.0
0 14   0.0 10.0  0.0
0 15  10.0 10.0  0.0
0 16   6.6  5.0  0.0
0 17   3.3 10.0  0.0
0 18  10.0  3.3  0.0
diffusity
TENSOR
3
MFACE
14
2 1    1.0 10.0 2.0
2 2    1.0 10.0 2.0
2 3    1.0 10.0 2.0
2 4    1.0 10.0 2.0
2 5    1.0 10.0 2.0
2 6    1.0 10.0 2.0
2 7    1.0 10.0 2.0
2 8    1.0 10.0 2.0
2 9    1.0 10.0 2.0
2 10   1.0 10.0 2.0
2 11   1.0 10.0 2.0
2 12   1.0 10.0 2.0
2 13   1.0 10.0 2.0
2 14   1.0 10.0 2.0
\end{verbatim}

\newpage
File in R1 format:
\begin{verbatim}
MSTK 1.0
R1 18 0 14 0
vertices
0.00 0.00 0.00  0 1
0.33 0.00 0.00  1 1
0.66 0.00 0.00  1 1
1.00 0.00 0.00  0 2
0.00 0.33 0.00  1 4
0.66 0.66 0.00  2 1
0.66 1.00 0.00  1 3
0.16 0.84 0.00  2 1
0.00 0.66 0.00  1 4
0.16 0.33 0.00  2 1
0.50 0.33 0.00  2 1
0.40 0.84 0.00  2 1
1.00 0.66 0.00  1 2
0.00 1.00 0.00  0 4
1.00 1.00 0.00  0 3
0.66 0.50 0.00  2 1
0.33 1.00 0.00  1 3
1.00 0.33 0.00  1 2
faces vertex
4  3 4 18 11              2 1
6  9 10 11 16 12 8        2 1
3  14 8 17                2 1
3  8 12 17                2 1
3  2 3 11                 2 1
4  1 2 10 5               2 1
4  6 16 18 13             2 1
4  6 13 15 7              2 1
3  7 17 12                2 1
3  5 10 9                 2 1
3  2 11 10                2 1
4  12 16 6 7              2 1
3  9 8 14                 2 1
3  11 18 16               2 1
attributes
velocity
VECTOR
3
MVERTEX
18
0 1    0.0  0.0  0.0
0 2    3.3  0.0  0.0
0 3    6.6  0.0  0.0
0 4   10.0  0.0  0.0
0 5    0.0  3.3  0.0
0 6    6.6  6.6  0.0
0 7    6.6 10.0  0.0
0 8    1.6  8.4  0.0
0 9    0.0  6.6  0.0
0 10   1.6  3.3  0.0
0 11   5.0  3.3  0.0
0 12   4.0  6.6  0.0
0 13  10.0  6.6  0.0
0 14   0.0 10.0  0.0
0 15  10.0 10.0  0.0
0 16   6.6  5.0  0.0
0 17   3.3 10.0  0.0
0 18  10.0  3.3  0.0
diffusity
TENSOR
3
MFACE
14
2 1    1.0 10.0 2.0
2 2    1.0 10.0 2.0
2 3    1.0 10.0 2.0
2 4    1.0 10.0 2.0
2 5    1.0 10.0 2.0
2 6    1.0 10.0 2.0
2 7    1.0 10.0 2.0
2 8    1.0 10.0 2.0
2 9    1.0 10.0 2.0
2 10   1.0 10.0 2.0
2 11   1.0 10.0 2.0
2 12   1.0 10.0 2.0
2 13   1.0 10.0 2.0
2 14   1.0 10.0 2.0
\end{verbatim}

@@

 if (argc == 1) {
   fprintf(stderr,"Usage: %s meshfilename (without .mstk extension)\n",argv[0]);
   exit(-1);
 }

@@

 mesh = MESH_New(UNKNOWN_REP);
 ok = MESH_InitFromFile(mesh,meshname);
 if (!ok) {
   fprintf(stderr,"Cannot find input file %s\n\n\n",meshname);
   exit(-1);
 }

@@

 if ((natt = MESH_Num_Attribs(mesh))) {

   fprintf(stderr,"Attributes on the mesh:\n\n");

   for (i = 0; i < natt; i++) {

     attrib = MESH_Attrib(mesh,i);

     attentdim = MAttrib_Get_EntDim(attrib);

     /* Won't print out edge, face and region based attributes
        but the code should be quite similar */

     if (attentdim != MVERTEX) continue;

     fprintf(stderr,"Attribute Number %-d\n",i+1);

     MAttrib_Get_Name(attrib,attname);
     fprintf(stderr,"Name: %-s\n",attname);

     atttype = MAttrib_Get_Type(attrib);
     switch(atttype) {
     case INT:
       fprintf(stderr,"Type: Integer\n");
       break;
     case DOUBLE:
       fprintf(stderr,"Type: Double\n");
       break;
     case VECTOR:
       fprintf(stderr,"Type: Vector\n");
       break;
     case TENSOR:
       fprintf(stderr,"Type: Tensor\n");
       break;
     default:
       fprintf(stderr,"Unrecognizable or unprintable attribute type\n");
       continue;	
     }

     switch(attentdim) {
     case MVERTEX:
       fprintf(stderr,"Applicable to vertices only\n");
       break;
     case MEDGE:
     case MFACE:
     case MREGION:
     case MALLTYPE:
       /* Won't print out edge, face and region based attributes
          but the code should be quite similar */
       continue; 
     default:
       fprintf(stderr,"Unrecognized entity type\n");
       continue;
     }

     ncomp = MAttrib_Get_NumComps(attrib);
     fprintf(stderr,"Number of components: %-d\n",ncomp);
